\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2016-11-17 redaction/dhp_sept2016.tex
changements dans le texte

\paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs. To do so, we start from the equation that expresses the price $p$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:

\begin{equation}
p = \tau \widetilde{p} + t \label{eq:base}
\end{equation}

\noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).  Let us denote $i$ the origin country, and $k$, the sector (or product). Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
\end{equation*}

@@

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
\end{equation*}

The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$). Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
\end{equation*}
\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Taking the logs, this drives us to estimate the following equation:

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= \ln(\varepsilon_{ik})$.

The non-linearity of Equation (\ref{eq:estimatedequation}) implies that it cannot be estimated using standard linear estimators. Therefore, all estimates are performed using non-linear least squares. The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares does not decrease from one iteration to the next.\todo{Dire plutot: sum of squares of residuals; to increase rather than to decrease? Not clear to me.} Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
These cut-offs are aimed at eliminating reporting or coding errors.\medskip

@@

\begin{eqnarray}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add} \\
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{eqnarray}

@@

\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs }  \\ \hline
    Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
    Median & 1.051 & 1.042 & 1.052 & 1.037 \\
    Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \textit{Ad-valorem term} & & & & \\ \hline
    Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
    Median & 1.028 & 1.018 & 1.028 & 1.016 \\
    Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term }& & & &   \\ \hline
    Mean  & 0.029 & 0.018 & 0.028 & 0.019 \\
    Median & 0.019 & 0.007 & 0.017 & 0.008 \\
    Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value. The additive term is expressed in fraction of fas price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{table}%

Table \ref{tab:summary_results} yields the following comments. First, the magnitude of overall transport costs is sizeable. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Air transport costs are smaller than sea transport costs. This is counter-intuitive, but certainly linked to selection effects.
Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable. Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%. These values are respectively equal to 1.8\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.

We investigate the robustness of this conclusion on a yearly basis. Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_with_add}), by transport mode.

@@

From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds on a yearly basis. For Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping. Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air transport. To get a better picture on this, we study the shares of both ad-valorem and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.

@@

At this stage, two interesting comments can be made. First, the magnitude of transport costs is higher for ocean shipping than for air transport. This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period. Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs. We come back to this aspect in Section \ref{sec:results_trends}.}

@@

In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fas prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs. To do so, we rely on several standard measures of fit. The first indicator is through comparing $R^{2}$. However, its use is far from being straightforward when evaluating non-linear estimates.\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one. In a non-linear context, $R^2$ is  strictly speaking inappropriate. However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.} As a complement, we provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line. Smaller values are better because it indicates that the observations are closer to the fitted line. We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test. A decrease in the log-likelihood function points to a better quality-of-fit. However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic, so that AIC stat $= 2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $. Once again, the preferred model is the one with the minimum AIC value. Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e. Equation (\ref{eq:model_with_add}) and the Restricted one (\emph{R}, i.e. Equation (\ref{eq:model_nlI})). The null tested is that the two models are statistically equivalent. Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level). The results for all years are available upon request to the authors.}\medskip

@@

Going deeper into the examination of the levels of these goodness-of-fit statistics reveals interesting differences. First, iceberg costs have more explanative power in the Vessel sector, where they account for 39\% of the variance of the CIF/FAB ratio, versus 31\% for the Air sector (considering the mean values of the $R^2$ across transport modes). The standard error of the regression based only on the ad-valorem term is consistently lower for Vessel than for Air.

Second, the dynamics over time also reveals that the quality of fit decreases strongly after 2000 for both transport modes. Interestingly, this seems to be attributable to different reasons depending of the transport mode considered. For Air, the additive term explains less and less variance over time: In 2000, the inclusion of the additive term increases the explanatory power by 30\%, and decreases the standard error of the regression by 14 percentage points;  in 2010 and 2013, it is hardly more than 10\% for the $R^{2}$, and 7 percentage points for the SER. Accordingly, the lower quality of fit for air transport seems to be attributable to a loss of explanatory power of the additive component. The picture is quite different for Vessel. In this case, the inclusion of the additive term increases the R$^{2}$ by a percentage very similar whatever the considered year (between 12 and 18\%); a similar conclusion can be drawn for the SER, which decreases by 8-11 percentage points across the years when additive costs are included. In contrast to air transport, the deteriorating performance of the model in terms of goodness-of-fit from 2000 mainly comes from the decreasing performance of the ad-valorem term. This suggests contrasting and time-evolving roles for additive and ad-valorem components of trade costs. We henceforth go deeper into this time-trend analysis in the next section.\smallskip

@@

\begin{eqnarray}
\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\Leftrightarrow \ln(\tau_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}

As for the additive cost, we estimate the following equation, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
%%In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
Equation (\ref{eq:compeffects_mult}) is estimated using OLS and Equation (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).

@@

@@

@@

\section{Robustness checks \label{app:robust_4d}}

>>>>

\paragraph{The estimated equation} Our purpose is to provide estimates over time of the size of ad-valorem and additive costs among total transport costs. To do so, we start from the equation that expresses the price $p(z)$ of a good $z$ paid by the importer (import, or cif) a function of the producer price $\widetilde{p}(z)$ (export, or fas), given both per-kg ($t$) and ad-valorem ($\tau$) transport costs, according to:

\begin{equation}
p(z) = \tau \widetilde{p}(z) + t \label{eq:base}
\end{equation}

\noindent As usual in the literature, the iceberg trade costs are denoted $\tau$ (with  $\tau \geq 1$, $\tau=1$ meaning no ad-valorem trade costs), while additive trade costs are labeled $t$ (with $t \geq 0$, $t=0$ implying no additive costs).  Let us denote $i$ the origin country, and $k$, the sector. Transforming the above equation (\ref{eq:base}) as ratio, we thus get the following equation at the root of our estimation:\footnote{We skip the year and transport-mode dimensions in the notations for reading convenience.}

\begin{equation*}
\frac{p_{ikz}}{\widetilde{p}_{ikz}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ikz}} \label{eq:base_estimee}
\end{equation*}

@@

\begin{equation*}
\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 =\tau_{i} \times \tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ikz}} \label{eq:theory_equation}
\end{equation*}

The ratio $\frac{p_{ikz}}{\widetilde{p}_{ikz}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ikz}>\widetilde{p}_{ikz}$). Taking into account this constraint in the estimation suggests that the error term should be always positive and multiplicative, as in:

\begin{equation*}
\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}} \right)\times \exp(\epsilon_{ikz})
\end{equation*}
\noindent where $\epsilon_{ik}$ follows an normal law centered on 0. Taking the logs, this drives us to estimate the following equation:

\begin{equation}
\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}}-1 \right) + \epsilon_{ikz} \label{eq:estimatedequation}
\end{equation}
where $\tau_{i}$, $\tau_{k}$, $t_{i}$ and $t_{k}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $k$, and $\epsilon_{ik}= \ln(\varepsilon_{ik})$.

The non-linearity of Equation (\ref{eq:estimatedequation}) implies that it cannot be estimated using standard linear estimators. Therefore, all estimates are performed using non-linear least squares. The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares does not increase from one iteration to the next. Finally, to eliminate the potential influence of outliers, we excluded 5 percent of the upper and lower tails of the distribution in the regression variables.
These cut-offs are aimed at eliminating reporting or coding errors.\medskip

@@

\begin{eqnarray}
\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ikz}}\right) + \epsilon_{ikz} \label{eq:model_with_add} \\
\ln\left(\frac{p_{ikz}}{\widetilde{p}_{ikz}}-1 \right)&=& \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ikz} \label{eq:model_nlI}
\end{eqnarray}

@@

\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    \multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air ($^{\ast \ast}$) & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs }  \\ \hline
    Mean  & 1.058 & 1.051 & 1.060 & 1.049 \\
    Median & 1.051 & 1.042 & 1.052 & 1.037 \\
    Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \textit{Ad-valorem term} & & & & \\ \hline
    Mean  & 1.032 & 1.025 & 1.033 & 1.024 \\
    Median & 1.028 & 1.018 & 1.028 & 1.016 \\
    Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \textit{Additive term }& & & &   \\ \hline
    Mean  & 0.029 & 0.018 & 0.028 & 0.019 \\
    Median & 0.019 & 0.007 & 0.017 & 0.008 \\
    Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \multicolumn{5}{l}{Data } \\ \hline
        Mean & 1.053 & 1.05& 1.056&1.039 \\
        Median & 1.043 & 1.02 & 1.044& 1.019 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value. The additive term is expressed in fraction of fas price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{table}%

Table \ref{tab:summary_results} yields the following comments. First, the magnitude of overall transport costs is sizeable. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Air transport costs are smaller than sea transport costs. This is counter-intuitive, but certainly linked to selection effects.
Second, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable. Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%. These values are respectively equal to 1.8\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}
We investigate the robustness of this conclusion on a yearly basis. Figure \ref{fig:mult_alone_withadd} thus reports the mean value of the ad-valorem component, estimated alone (Dashed line, corresponding to Equation (\ref{eq:model_nlI})) and along with the mean value of the additive component (Plain line, corresponding to Equation (\ref{eq:model_with_add}), by transport mode.

@@

From Figure \ref{fig:mult_alone_withadd}, we can infer that the sizeable magnitude of additive costs also holds on a yearly basis. Further, Figure \ref{fig:mult_alone_withadd} reports that for Air, the size of the bias seems to decrease over time, while it appears rather stationary for ocean shipping. Put it differently, this suggests that the share of the additive component in total transport costs has decreased over time for air transport. To get a better picture on this, we study the shares of both ad-valorem and additive components in total costs, by transport mode and over time, based on Figure \ref{fig:decomp_TC_3d}.

@@

At this stage, two interesting comments can be made. First, the magnitude of transport costs is higher for ocean shipping than for air transport. This result holds considering ad-valorem costs alone (Figure \ref{fig:mult_alone_withadd}), as well as in presence of costs (Figure \ref{fig:decomp_TC_3d}), for all years throughout the period. Second, in both transport modes, the additive component appears of sizeable importance: As mean value over the period, it amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. We go on investigating this result further in the next subsection.\footnote{Figure \ref{fig:decomp_TC_3d} also delivers interesting results with respect to the time trends of transport costs. We come back to this aspect in Section \ref{sec:results_trends}.}\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}

@@

In this section, we explore the performances of each type of model (with and without additive costs) in fitting the observed cif-fas prices gap, in order to deliver a more systematic diagnosis about the importance of additive costs. To do so, we rely on several standard measures of fit. The first indicator is through comparing $R^{2}$. However, its use is far from being straightforward when evaluating non-linear estimates.\footnote{$R^2$ is based on the underlying assumption that the adjusted model is a linear one. In a non-linear context, $R^2$ is  strictly speaking inappropriate. However, if the error distribution is approximately normal, a standard metric like $R^2$ remains informative on the quality of adjustment.} This drives us to complement the goodness of fit diagnosis with three alternative measures. We provide the Standard Error of Regression (SER), which represents the average distance that the observed values fall from the regression line. Smaller values are better because it indicates that the observations are closer to the fitted line. We also report the log-likelihood function, and two measures derived, the Akaike Information Criterion (AIC) and the log-likelihood (LL) ratio test. A decrease in the log-likelihood function points to a better quality-of-fit. However, the likelihood function systematically decreases with the number of parameters included; the AIC criterion allows for correcting this overfitting by including a penalty in the computation of the statistic.\footnote{Precisely, the AIC stat is equal to $2 \times \textrm{number of parameters} - 2 \times \textrm{Likelihood} $.} The preferred model is the one with the minimum AIC value. Finally, the log-likelihood ratio test statistic compares systematically the likelihood of the Unrestricted model (\emph{UR}, including an additive term, i.e. Equation (\ref{eq:model_with_add}) and the Restricted one (\emph{R}, i.e. Equation (\ref{eq:model_nlI})). The null tested is that the two models are statistically equivalent. Results are reported in Tables \ref{tab:good_fit_air} and \ref{tab:good_fit_vessel}, for Air and Vessel respectively, at the 3-digit level.\footnote{Due to our time coverage, we do not report the results for all years (at the 3-digit level). The results for all years are available upon request to the authors.}\medskip

@@

\smallskip

@@

\begin{eqnarray}
\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\Leftrightarrow \ln(\widehat{\tau}_{ikt})&=&\delta +\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i + \sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}

As for the additive cost, we estimate the following equation, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\widehat{t}_{ikt}&=&\left( \sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\sum_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\Leftrightarrow \ln(t_{ijt})&=&\ln\left( \sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\sum_{k}\beta_k.\mathbb{1}_k\right) + \sum_{t \neq 1974}\gamma_t.\mathbb{1}_t+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
Notice that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions.
%%In the three estimations (ad-valorem costs alone, ad-valorem costs estimated along with additive, and the additive component), we consider Argentina, the sector 011 and the first year of our dataset 1974, as references for the country-, product and -year dummies.
Equation (\ref{eq:compeffects_mult}) is estimated using OLS and Equation (\ref{eq:compeffects_add}) is estimated using non-linear least squares (by transport mode).

@@

\textbf{Faire l'analyse mais sur l'ecart cif-fob observe et non plus reconstruit. Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}

@@

\section{Robustness checks}

\subsection{Robustness to the specification of the estimated equation}

\paragraph{Estimating the difference between fas and cif prices in level}

Ie, estimate

$$\log(p_{ikz})= \log(\tau_i\times \tau_k \times \widetilde{p}_{ikz}+ t_i+ t_k)+ \exp(\epsilon_{ikz})$$

\paragraph{Robustness to the separability assumption}

Ie, estimate

$$\log(\frac{p_{ikz}}{\widetilde{p}_{ikz}} -1)= \log(\tau_{ik} -1+ \frac{t_{ik}}{\widetilde{p}_{ikz}})+ \exp(\epsilon_{ikz})$$

@@

\subsection{Goodness-of-fit: Detailed results \label{app:robust_4d}}
